Gravitational Lens Recovery with \Glass: Measuring the mass profile and shape of a lens

### 2014-01-23 limits/ms.tex
Added shape paragraph and short conclusion. Acknowledgments.

The pixels only cover a finite circular area with physical radius $\Rmap$ and
pixel radius $\Rpix$ with the central cell centered on the lensing galaxy. To account for any global
shearing outside this region from, e.g., a neighboring galaxy, we also add to
\eqnref{discrete potential} two shearing terms
%
\begin{equation}
\label{shear}
\gamma_1(\theta_x^2 - \theta_y^2) + 2\gamma_2\theta_x\theta_y\quad.
\end{equation}
%
We can continue adding terms to account for other potentials. For instance,
we may want to impose a base potential over the field, or add potentials
from the presence of other galaxies in the field. \Glass\
already includes potentials for a point mass or an exponential form, but custom
potentials are straightforward to add and can be included directly in the input file.
If the stellar density $\kappa_s$ has been estimated we can use this as a lower bound
where the stellar potential is a known constant of the form \eqnref{discrete
potential}, e.g., $\kappa_n = \kappa_\mathrm{DM} + \kappa_s$ for a two-component model.

@@

The main results from modeling these different configurations are shown in
\figref{main results} and \figref{main results pixel-wise}. Each
subplot corresponds to a different mock galaxy. We show the fractional error of the mass
distribution for each of the test configurations with (red) and without stellar mass (black)
In \figref{main results}, the error is defined over radial bins of mass as
%
\begin{equation} \label{ferror R}
  f_R = \frac {\sum_i \left|M_p(i) - \tilde M_p(i)\right| } {\sum \tilde M_p(i)}
\end{equation}
%
where $M_p$ is the mass in a radial bin and $\tilde M_p$ is the mass from the mock galaxy.
In \figref{main results pixel-wise} the error is defined over all the pixels $\vec\theta$
%
\begin{equation} \label{ferror theta}
f_\theta = \frac {\sum_{\vec\theta} \left|M(\vec\theta) - \tilde M(\vec\theta)\right| } {\sum \tilde M(\vec\theta)}
\end{equation}
%
%\begin{equation}
%  \chi_i \equiv 100 \times \sqrt{\frac{\int_r (\rho_{\M_i}(r) - \rho_G(r))^2}{\int_r \rho_G(r)^2}}
%\chi_i = \frac{\sum_r (\rho(r)_{\M_i} - \rho(r)_G)^2}{\sum_r \rho(r)_G^2}
%\end{equation}
%where $\rho_{\M_i}(r)$ is the recovered density profile for a mock model $\M_i$; and $\rho_G(r)$ is the true density profile.

@@

@@

The abundance of strong lensing data increases from left to right within each plot. As
a result, there is a general trend for the reconstruction quality to increase
(and for $f$ to decrease). By adding more measurable information to each
configuration the quality can also be affected. When both time delays and a
central image are present the quality is highest. A double is known to provide
very little constraint on the mass distribution. This is particularly evident in
galaxies \mockAC{} and \mockBC{} where the mass profile is steepest and the reconstruction of
the double is poorest. However, the addition of an arc is sufficient to correct
this.

\hilight{Write about the shape plot and how we measured shape.}

\begin{figure*}
\includegraphics[width=0.33\textwidth]{BCQuadR1a_Tms_sigp-1.pdf}
\includegraphics[width=0.33\textwidth]{BCQuadR1a_TmS_sigp-2.pdf}
\includegraphics[width=0.33\textwidth]{BC_beta.pdf}
\caption{ \hilight{Larger fonts.} Estimated projected radially averaged velocity dispersion $\sigma_p$
\eqnrefp{eqn:sphericaljeans} for two \mockBC{} models ZContrastR1R3
(\textbf{left}) and QuadR1a (\textbf{middle}) assuming an anisotropy $\beta=0$
(blue, solid) and $\beta=1$ (blue, dashed).  The equivalent
curves are also shown for the projected mock data after using the same analysis routines
(purple). The solid black line is the actual cylindrically averaged velocity
dispersion of the original mock particle data.  The stellar half mass radius
(yellow) Einstein radius (black) are marked by vertical lines. For this
configuration, these two radii are well-separated. The actual variation in $\beta(r)$ is show also shown (\textbf{right}).}

@@

With the stellar mass lower bound, the improvement of the reconstruction
quality shown in \figref{main results} is quite dramatic for the
steepest mock galaxies (AC and BC). This is because these models are dominated
by stars in the inner region. By contrast, the other two galaxies -- where the stars contribute negligibly to the potential -- are unaffected.

@@

As outlined in \S\ref{sec:glass}, \Glass\ can also run post processing routines
on the model ensemble which can be used to apply non-linear constraints. As an
example, we consider here constraints from stellar kinematics. The models in
the \Glass\ ensemble are processed as described in \S\ref{sec:glasskinematics}.
To illustrate the power of stellar kinematic constraints, in
\figref{sigma-beta}, we plot the projected velocity dispersion calculated for
one model model (extracted from the full ensemble) of the BC Quad with time
delays and stellar mass (left), and the mean projected velocity dispersion
\eqnrefp{eqn:virialaverage} as a function of the upper limit of the averaging
integral (the aperture size), $a_p$ (right). In both cases, we calculate curves
for two extrema velocity anisotropies: $\beta=0$ (green) and $\beta=1$ (red).
Over-plotted is the correct answer for the BC model (black). The stellar half
mass radius (yellow) Einstein radius (black) are marked by vertical lines. For
this configuration, these two radii are well-separated.

@@

\hilight{This paragraph should go.}
We now consider formally accepting/rejecting models from the full ensemble. We
consider two regimes of interest: a model with a double where the mass profile
is poorly constrained \hilight{(XXX)}, and a model where the mass profile is very well
constrained by the lens data \hilight{(YYY)}. In the former case, we use
$\overline{\sigma}(2 r_{1/2})$ to probe the mass independently of $\beta(r)$.
We accept/reject models assuming a Gaussian error on measured
$\overline{\sigma}(2 r_{1/2})$ of $\pm 10$\,km/s (similarly to the data quality
presented for example in \citealt{2002MNRAS.337L...6T}). The results are shown
in \figref{main results}, blue data points. In the latter case, we ask whether
the data favour $\beta = 0$ or $\beta = 1$ by obtaining a $\chi^2$ goodness of
fit measure between the true and modelled $\sigma_p(R)$. We accept/reject
models from the ensemble above. The results for $\beta$ are shown in Figure
ZZZ.

The results for stellar kinematics match our expectations from
\S\ref{sec:kinematics}. Where the lens data already constrain the mass
distribution at $r \sim r_{1/2}$, stellar kinematics provide valuable
information about the velocity anisotropy of the stars, $\beta$ (see Figure
ZZZ). Where the lens data poorly constrain the mass distribution at $r_{1/2}$,
we may `integrate out' the effect of unknown $\beta$ to obtain a robust measure
of $M(r_{1/2})$ from the stellar kinematics. This latter is robust to both
uncertainties in $\beta(r)$ and to our assumption of spherical symmetry in the
kinematic models \citep{2012ApJ...754L..39A}. Finally, we note that there is a
third interesting case. If the lens models well-constrain $M(r_{1/2})$ and
stellar kinematics are available, then these can be used to probe cosmological
models (see \S\ref{sec:theory}). We explore this in more detail in a separate
publication.

@@

Various action items.

>>>>

The pixels only cover a finite circular area with physical radius $\Rmap$ and
pixel radius $\Rpix$ with the central cell centered on the lensing galaxy. To account for any global
shearing outside this region from, e.g., a neighboring galaxy, we also add to
\eqnref{discrete potential} two shearing terms
%
\begin{equation}
\label{shear}
\gamma_1(\theta_x^2 - \theta_y^2) + 2\gamma_2\theta_x\theta_y\quad.
\end{equation}
%
We can continue adding terms to account for other potentials. For instance,
we may want to impose a base potential over the field, or add potentials
from the presence of other galaxies in the field. \Glass\
already includes potentials for a point mass or an exponential form, but custom
potentials are straightforward to add and can be included directly in the input file.
If the stellar density $\kappa_s$ has been estimated we can use this as a lower bound
where the stellar potential is a known constant of the form \eqnref{discrete
potential}, e.g., $\kappa_n = \kappa_{\mathrm{dm},n} + \kappa_{s,n}$ for a two-component model.

@@

The main results from modeling these different configurations are shown in
\figref{main results} and \figref{main results pixel-wise}. Each
subplot corresponds to a different mock galaxy. We show the fractional error of the mass
distribution for each of the test configurations with (red) and without stellar mass (black)
In \figref{main results}, the error is defined over radial bins of mass as
%
\begin{equation} \label{ferror R}
  f_R = \frac {\sum_i \left|M(i) - \tilde M(i)\right| } {\sum \tilde M(i)}
\end{equation}
%
where $M$ is the mass in bin $i$ and $\tilde M$ is the mass from the mock galaxy.
In \figref{main results pixel-wise} the error is defined over all the pixels $\vec\theta$
%
\begin{equation} \label{ferror theta}
f_\theta = \frac {\sum_{\vec\theta} \left|M(\vec\theta) - \tilde M(\vec\theta)\right| } {\sum_{\vec\theta} \tilde M(\vec\theta)}
\end{equation}
%
%\begin{equation}
%  \chi_i \equiv 100 \times \sqrt{\frac{\int_r (\rho_{\M_i}(r) - \rho_G(r))^2}{\int_r \rho_G(r)^2}}
%\chi_i = \frac{\sum_r (\rho(r)_{\M_i} - \rho(r)_G)^2}{\sum_r \rho(r)_G^2}
%\end{equation}
%where $\rho_{\M_i}(r)$ is the recovered density profile for a mock model $\M_i$; and $\rho_G(r)$ is the true density profile.

@@

With both $f_R$ and $f_\theta$ we only consider mass up to one radial
position passed the outermost image, since there is no longer any lensing
information beyond that point.

The abundance of strong lensing data increases from left to right within each
plot. As a result, there is a general trend for the reconstruction quality to
increase (and for $f$ to decrease). By adding more measurable information to
each configuration the quality can also be affected. When both time delays and
a central image are present the quality is highest. A double is known to
provide very little constraint on the mass distribution. This is particularly
evident in galaxies \mockAC{} and \mockBC{} where the mass profile is steepest
and the reconstruction of the double is poorest. However, the addition of an
arc from the extended source is sufficient to correct this. \hilight{Mention the saturation of the two panels.}

We can also directly measure the recoverability of the shape of the mass
distribution by computing the eigenvalues of the inertia tensor. We define our
shape by the ratio of these eigenvalues $s=\lambda_1/\lambda_2$, where the
$\lambda_1$ is always the largest value. Again, we only consider mass up to one
radial position passed the outermost image and compute our the fractional error
%
\begin{equation} \label{ferror shape}
  f_\mathrm{shape} = \left|s - \tilde s\right|  / \tilde s
\end{equation}
%
where $\tilde s$ is the shape of the mock galaxy. The distribution of $f_\mathrm{shape}$
for each mock galaxy and each test case is shown in \figref{shape results}.

@@

\begin{figure*}
\includegraphics[width=0.33\textwidth]{BCQuadR1a_Tms_sigp-1.pdf}
\includegraphics[width=0.33\textwidth]{BCQuadR1a_TmS_sigp-2.pdf}
\includegraphics[width=0.33\textwidth]{BC_beta.pdf}
\caption{ \hilight{Larger fonts.} Estimated projected radially averaged velocity dispersion $\sigma_p$
\eqnrefp{eqn:sphericaljeans} for a single quad from the \mockBC{} mock galaxy without stellar mass
(\textbf{left}) and with stellar mass (\textbf{middle}) assuming an anisotropy $\beta=0$
(blue, solid) and $\beta=1$ (blue, dashed).  The equivalent
curves are also shown for the projected mock data after using the same analysis routines
(purple). The solid black line is the actual cylindrically averaged velocity
dispersion of the original mock particle data.  The stellar half mass radius
(yellow) Einstein radius (black) are marked by vertical lines. For this
configuration, these two radii are well-separated. The actual variation in $\beta(r)$ is show also shown (\textbf{right}).}

@@

\hilight{This is a bit overstated.}
With the stellar mass lower bound, the improvement of the reconstruction
quality shown in \figref{main results} is quite dramatic for the doubles in the 
steepest mock galaxies (\mockAC{} and \mockBC). This is because these models are dominated
by stars in the inner region. By contrast, the other two galaxies -- where the stars contribute negligibly to the potential -- are unaffected.

@@

As outlined in \S\ref{sec:glass}, \Glass\ can also run post processing routines
on the model ensemble which can be used to apply non-linear constraints. As an
example, we consider here constraints from stellar kinematics. The models in
the \Glass\ ensemble are processed as described in \S\ref{sec:glasskinematics}.
To illustrate the power of stellar kinematic constraints, in
\figref{fig:sigp}, we plot the projected velocity dispersion calculated for
one model model (extracted from the full ensemble) of the \mockBC{} Quad with time
delays and stellar mass (left), and the same but without stellar mass (right).
In both cases, we calculate curves
for two extrema velocity anisotropies: $\beta=0$ (green) and $\beta=1$ (red).
Over-plotted is the correct answer for the \mockBC{} model (black). The stellar half
mass radius (yellow) Einstein radius (black) are marked by vertical lines. For
this configuration, these two radii are well-separated.

@@

\hilight{This paragraph should be removed.}
We now consider formally accepting/rejecting models from the full ensemble. We
consider two regimes of interest: a model with a double where the mass profile
is poorly constrained \hilight{(XXX)}, and a model where the mass profile is very well
constrained by the lens data \hilight{(YYY)}. In the former case, we use
$\overline{\sigma}(2 r_{1/2})$ to probe the mass independently of $\beta(r)$.
We accept/reject models assuming a Gaussian error on measured
$\overline{\sigma}(2 r_{1/2})$ of $\pm 10$\,km/s (similarly to the data quality
presented for example in \citealt{2002MNRAS.337L...6T}). The results are shown
in \figref{main results}, blue data points. In the latter case, we ask whether
the data favour $\beta = 0$ or $\beta = 1$ by obtaining a $\chi^2$ goodness of
fit measure between the true and modelled $\sigma_p(R)$. We accept/reject
models from the ensemble above. The results for $\beta$ are shown in the right panel of
\figref{fig:sigp}.

The results for stellar kinematics match our expectations from
\S\ref{sec:kinematics}. Where the lens data already constrain the mass
distribution at $r \sim r_{1/2}$, stellar kinematics provide valuable
information about the velocity anisotropy of the stars, $\beta$ (see \figref{fig:sigp}). Where the lens data poorly constrain the mass distribution at $r_{1/2}$,
we may `integrate out' the effect of unknown $\beta$ to obtain a robust measure
of $M(r_{1/2})$ from the stellar kinematics. This latter is robust to both
uncertainties in $\beta(r)$ and to our assumption of spherical symmetry in the
kinematic models \citep{2012ApJ...754L..39A}. Finally, we note that there is a
third interesting case. If the lens models well-constrain $M(r_{1/2})$ and
stellar kinematics are available, then these can be used to probe cosmological
models (see \S\ref{sec:theory}). We explore this in more detail in a separate
publication.

@@

We have introduced a new gravitational len modelling tool---\Glass---and used
it to test the recovery of the mass profile and shape of mock strong lensing
galaxies. We have also shown that with the addition of stellar mass
constraints, one can better constrain the inner profile of galaxies,
particularly in the case of doubles. Furthermore, by we have demonstrated our
sensitivity to the estimates of the velocity dispersion. Improved measurements
of the dispersion and, in particular, the velocity anisotropy could be used to
greatly enhance our lens models.

\section{Acknowledgments}\label{sec:Acknowledgements}

The authors would like to thank Sarah Bryan and Walter Dehnen for creating the
particle distributions for the mock galaxies. This work has been partially
funded by \hilight{if we want..}
